Detecting Facial Expressions in Video Stills from Professional Tennis Matches

### 2017-12-15 paper/paper.Rmd

This paper examines the effectiveness of facial detection APIs on broadcast video stills of Australian Open tennis matches. The goal is to determine the best API to use for face detection of players throughout a match. For training purposes faces were manually tagged in 6406 images, recording the scene characteristics and features of individual faces. This included information regarding accessories, such as headwear or sunglasses being worn. This enables the performance to be assessed based on conditions, and the APIs were evaluated on their success rate at detecting these faces. The possible obstacles of real time implementation were also assessed via the time taken to complete detection within an image.

@@

\begin{table}[htp]
\centering
\caption{Table of the amount and proportion of faces each API captured, by manual tagging.}
\label{tab:facecounts}
\begin{tabular}{lllll}\\\cline{1-5}
            & \multicolumn{2}{l}{Faces found} & \multicolumn{2}{l}{Non-faces detected as faces} \\\cline{1-5}

            & Count        & Proportion       & Count                & Proportion               \\\cline{1-5}
Google      & 1319         & 0.57             & 638                  & 0.49                     \\
Microsoft   & 565          & 0.24             & 505                  & 0.38                     \\
Skybiometry & 493          & 0.22             & 512                  & 0.39                     \\
Animetrics  & 434          & 0.19             & 527                  & 0.40\\ \cline{1-5}
\end{tabular}
\end{table}

The mismatches were manually examined, and showed that the `r amin %>% filter(type=="Google",!matchesManual) %>% nrow()` potential faces found by Google but not tagged manually were actually faces, but were crowd members. They would need to be used, for the long term purpose of the study, to detect and tag players' emotions during a match.

Comparatively, visual inspection of the `r amin %>% filter(type=="Animetrics",!matchesManual) %>% nrow()` potential faces that did not match manually annotated faces showed the Animetrics results contained many potential faces that were unusual areas to be considered faces, this is discussed in the false discoveries section.

```{r FaceUpSet, fig.cap = "\\label{fig:FaceUpset} Faces Per API Combination. The Bar Chart shows the faces that were recognised by multiple API or found manually. The largest group, with 946 faces, is faces only found by Manual annotations. The following group were the 716 faces recognized both Manually and by the Google API. These combinations give insight to the algorithm specifications as some APIs certain faces when others do not."}
createUpSet(aminFaces)
```

@@

The colours in the mosaic \@ref(fig:ICFigs) (a) in Figure \@ref(fig:ICFigs) show the proportion of faces that matched those found manually, given the angle the faces were captured from. There is a greater proportion of API faces that matched the faces found manually than those that did not match. However given the face was captured from a birds eye angle there were a lot less faces proportionally that matched those found manually. The most common faces in the set were those captured level to player's faces, these faces were often recognised by the APIs as well.

@@

This mosiac, Figure \@ref(fig:ICFigs) (d), allows consideration of the interaction between the background of the image captured and the situation that could possibly be occurring. As seen previously, the logo wall is the most common background, but it is never the background of an image of the crowd. The court is the background of an image only when players are captured, this occurs during close ups and while the court is in play.

@@

The use of accessories like glasses and headwear, visors or hats, was considered as the Australian Open takes place on both indoor and outdoor courts. 
It was assumed that outdoor courts would lead to the use of these accessories and these accessories may contribute to the performance of a recognition software.

@@

\begin{table}[ht]
\centering
%\parbox{.5\linewidth}{
\caption{Proportions of faces captured manually tagged as wearing headwear given they wore glasses, or did not.} 
\label{tab:accessories}
\begin{tabular}{ccc}
\hline 
 & Head wear & No Head wear\\ 
\hline
Glasses & 0.50 & 0.50 \\
No Glasses & 0.26 & 0.74 \\
\hline
\end{tabular}
\end{table}

Table \@ref(tab:accessories) of the proportions tells that of the faces wearing glasses, there were a similar amount of faces captured with and without headwear. Of those without glasses only 26% wore headwear. This may be influenced by weather and sunlight on the court during Australian Open matches.

>>>>

This paper examines the effectiveness of facial detection APIs on broadcast video stills of Australian Open tennis matches. The goal is to determine the best API to use for face detection of players throughout a match. For training purposes faces were manually tagged in 6406 images, recording the scene characteristics and features of individual faces. This included information regarding accessories, such as head wear or sunglasses being worn. This enables the performance to be assessed based on conditions, and the APIs were evaluated on their success rate at detecting these faces. The possible obstacles of real time implementation were also assessed via the time taken to complete detection within an image.

@@

\begin{table}[htp]
%\parbox{.5\linewidth}{
\caption{Proportion of API matches with the 2295 manually tagged faces, and faces detected in the 1314 images with no manually tagged faces.}
\bigskip
\label{tab:facecounts}
\centering
\begin{tabular}{|l|rr|rr|}  \hline
& \multicolumn{2}{r|}{Faces found}&\multicolumn{2}{r|}{Non-faces found}\\\cline{2-5}
API & Count & Prop & Count & Prop \\\hline
Google & 1319 & 0.57 & 638 & 0.49 \\
Microsoft & 565 & 0.25 & 505 & 0.38 \\
Skybiometry & 493 & 0.21 & 512 & 0.39 \\
Animetrics & 434 & 0.19 & 527 & 0.40 \\\hline
\end{tabular}
\end{table}

```{r FaceUpSet, fig.cap = "\\label{fig:FaceUpset} Set visualization, to examine intersections between API and manual face detections:  bar charts show counts, by API (lower left) and for combinations (main part of plot), and dot display (bottom) indicates the group intersections. All five way intersections are examined, but only the first few have large counts. The largest group is the manual tagging with 946 faces uniquely identified. The intersection between manual tagging and Google is 716 faces. The intersection of all four APIs and manual tagging is fourth highest, at 342."}
setdata <- aminFaces %>% 
  mutate(count = 1) %>%
  select(type, FaceID, count) %>%
  separate(FaceID, c("id1", "id2", "API")) %>%
  unite(id, c("id1", "id2")) %>%
  spread(API, count, fill=0) %>% 
  arrange(id) %>% 
  group_by(id) %>%
  summarise(Animetrics=as.integer(sum(An)), 
            Google=as.integer(sum(Go)), 
            Manual=as.integer(sum(Ma)),
            Microsoft=as.integer(sum(Mi)), 
            Skybiometry=as.integer(sum(Sk)))

setdata <- setdata %>% mutate(id=factor(id)) 
setdata <- as.data.frame(setdata)

upset(setdata, order.by="freq")
# colsum <- apply(setdata[,-1], 2, sum)
# rowsum <- apply(setdata[,-1], 1, sum)
# onlyone <- setdata[rowsum==1,] 
# apply(onlyone[,-1], 2, sum)
```

@@

The colors in the mosaic \@ref(fig:ICFigs) (a) in Figure \@ref(fig:ICFigs) show the proportion of faces that matched those found manually, given the angle the faces were captured from. There is a greater proportion of API faces that matched the faces found manually than those that did not match. However given the face was captured from a birds eye angle there were a lot less faces proportionally that matched those found manually. The most common faces in the set were those captured level to player's faces, these faces were often recognized by the APIs as well.

@@

This mosaic, Figure \@ref(fig:ICFigs) (d), allows consideration of the interaction between the background of the image captured and the situation that could possibly be occurring. As seen previously, the logo wall is the most common background, but it is never the background of an image of the crowd. The court is the background of an image only when players are captured, this occurs during close ups and while the court is in play.

@@

The use of accessories like glasses and head wear, visors or hats, was considered as the Australian Open takes place on both indoor and outdoor courts. 
It was assumed that outdoor courts would lead to the use of these accessories and these accessories may contribute to the performance of a recognition software.

@@

\begin{table}[ht]
\centering
%\parbox{.5\linewidth}{
\caption{Proportions of faces captured manually tagged as wearing headgear given they wore glasses, or did not.} 
\label{tab:accessories}
\begin{tabular}{ccc}
\hline 
 & Head wear & No Head wear\\ 
\hline
Glasses & 0.50 & 0.50 \\
No Glasses & 0.26 & 0.74 \\
\hline
\end{tabular}
\end{table}

Table \@ref(tab:accessories) of the proportions tells that of the faces wearing glasses, there were a similar amount of faces captured with and without head wear. Of those without glasses only 26% wore head wear. This may be influenced by weather and sunlight on the court during Australian Open matches.
